---
title: "DietTable and DNW"
author: "BH & JP"
date: '2017-06-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(httr)
library(plyr)
library(ggplot2)
library(mgcv)
library(reshape2)
library(gtools)

```




## Things to do

Read in fish DietTable form globi and fish ecology table from globi-fishbase
Cobine files and select relevant columns (i.e diet, habitat (marine/freshwater), trophic level)

NOTE - no consumers prey on chondricthyes or clittelata - these can be removed but not rightn now as that will mess up the code...


```{r}
ecology <- read.csv('ecology_fishbase.csv')
diet <- read.csv('fbDietMatrix20170607.tsv', sep='\t')
dietEcology <- merge(x = diet, y = ecology, by = 'SpecCode')

diet.sum<-dietEcology[,c(1:33, 63,68,73,74,79,80)]

```


## Convert diet to presence/absence

Replace diet values with 1/0


```{r}
diet.sum[,c(5:33)][diet.sum[,c(5:33)] > 0] <- 1
```


## Generate number of GBIF occurences and test relationship with DNW

The next step is to determine the number of occurence of each species in GBIF.

Dieatry niche width may be a function of data availability so need to test this before doing any subsequent analysis with DNW values

Script to retrieve GBIF occ by GBIF id - this takes 10-15 mins to run so avoid where possible
to save time I have saved the GBIF counts in a seperate DF, delete these lines from the final script...
```{r}

diet.sum$gbif.taxon.occ.count <- lapply(diet.sum$gbif.taxon.id, function(taxonKey) {
	resp <- httr::GET(paste('http://api.gbif.org/v1/occurrence/count?taxonKey=',taxonKey, sep=""))
	message('.', appendLF=FALSE)
	httr::stop_for_status(resp)
	httr::content(resp)
})

diet.sum$gbif.taxon.occ.count <- unlist(diet.sum$gbif.taxon.occ.count)
diet.sum<-subset(diet.sum, gbif.taxon.occ.count > 0)
diet.sum$OCClog <- log(diet.sum$gbif.taxon.occ.count)
GBIF<-diet.sum[,c(1:3,40)]

```


##Split diet table into marine and freshwater fishes

#Marine

subset to only fish recorded in marine habitats
Generate DNW measurement n prey types / total prey types (29)
Remove any species with DNW = 0

```{r}

marine<-subset(diet.sum, Marine == -1 & Freshwater == 0)
marine$DNW<-rowSums(marine[,c(5:8,10:26,28:33)])/27 #remove clitelata and chondrythes now...
marine<-subset(marine, DNW > 0)

```


Plot relationship between DNW and GBIF occurences, GBIF occurences are log transformed to normalise the range

```{r}
ggplot(marine, aes(OCClog, DNW)) +
  geom_point() +
  geom_smooth(method = "loess")

```
Positive relationship between DNW and OCClog
Run a gam to 'correct' for this relationship.
DNWcorr is the residuals of this model and will be used for subsequent analysis

```{r}
gam.marine<-gam(DNW ~ s(OCClog), data=marine)
summary(gam.marine)
marine$DNWcorr<-residuals(gam.marine)
marine$biome<-"Marine"

```

##Freshwater

subset to only get species observed in Freshwater

Several prey groups (namely; cnidaria.other, Anthozoa, Clitellata, Polyplacophora, Cephalopoda, Brachiopoda, Echinodermata, Ophiuroidea, Holothuroidea) are exclusively marine groups and were not recorded in the diet of freshwater fish. These groups were removed from DNW calculation. 

NOTE - Cnidaria.other, Cephalopoda & Echinodermata were prey categories for a small number (n<10) freshwater fishes. In some cases this is anadromous species (eg Onchyrinchus sp.) but others (eg cyprinids: Tinca tinca, Rutilus rutilus) seem strange. These may require some futher checking of GLOBI sources....

Next step is to caculate DNW for freshwater fishes, n prey types / total prey types (18)

```{r}
fresh<-subset(diet.sum, Marine == 0 & Freshwater == -1)
summary(fresh)
fresh<-fresh[,-c(5,6,9,11,14,15,21:24,27)]
fresh$DNW<-rowSums(fresh[,c(5:22)])/18
fresh<-subset(fresh, DNW > 0)


```



Plot relationship between DNW and GBIF occurences, GBIF occurences are log transformed to normalise the range

```{r}
ggplot(fresh, aes(OCClog, DNW)) +
  geom_point() +
  geom_smooth(method = "loess")

```
Positive relationship between DNW and OCClog
Run a gam to 'correct' for this relationship.
DNWcorr is the residuals of this model and will be used for subsequent analysis

```{r}
gam.fresh<-gam(DNW ~ s(OCClog), data=fresh)
summary(gam.fresh)
fresh$DNWcorr<-residuals(gam.fresh)
fresh$biome<-"Fresh"

```

## Combine marine and freshwater datasets as new DietTable

Combine marine and freshwater datasets
Quick plot to check everything and compare between biomes

```{r}

DietTable<-smartbind(marine, fresh)
write.csv(DietTable, "DietTable.csv")
ggplot(DietTable, aes(DNWcorr, fill=biome, colour=biome)) +
  geom_density(alpha=0.2) 



```
 
 Now for some fun comparisons....
 
 is there any relationship between niche width and trophic level?
 First need to change TL values from factor to numberic
 
```{r}
DietTable$DietTroph<-as.numeric(DietTable$DietTroph)
DietTable$FoodTroph<-as.numeric(DietTable$FoodTroph)
ggplot(DietTable, aes(y = DNWcorr, x = FoodTroph/100, colour=biome)) +
  geom_point(colour = "grey") +
  geom_smooth(method = "loess") +
  facet_grid(. ~ biome)


```




PCA to search for trends in diet - nothing obvious, likely that the opresence absence 

```{r}
mar.PC<-marine[,c(5:8,10:26,28:33)]
summary(DietTable)
marine.pca <- prcomp(mar.PC, scale = TRUE) 
biplot(marine.pca)
summary(marine.pca)


```